Does personal relevance attenuate citizens' persuasion by government branding? Evidence from a survey experiment on air-pollution policy

### 2019-07-23 paper3_APPENDIX_revision_01.Rmd
Add files via upload

@@

>>>>

 and restricting the sample to those who saw that policy first before the recycling policy. The latter analysis rules out the possibility that subjects' responses are affected by the manipulations in the first policy presented to them (a carryover effect), or by their evaluation of the it (assimilation or contrast effects).

Notice that the effect of the real symbols is greater, which may indicate that the impact of symbols diminishes between the first and the second time participants see them.

@@

The subsequent media reports specifically highlighted the conclusion that half of the cases of cancer among children in the area can be associated to air-pollution [@elroi_website_2015; @raavad_website_2015]. 
Some reports also highlighted that the areas that are geographically closer to the industrial areas are found to be at higher risk for cancer [@bendavid_website_2016]. 
Finally, as an additional test for the robustness of my findings, I further compare between Haifa Bay subjects who are more/less likely to perceive the air-pollution as personally relevant, based on two proxies. First, I rely on subjects city of residence, and its geographical proximity to the polluting industrial area. Congruently, I selected the three cities/suburbs that are closest to the industrial area - *Nesher*, *Qiryat Haiim* and *Qiryat Bialik* (*n*=`r SVIVA2_01_haifa %>% filter(close.industrial.area==1) %>% nrow()`) -- and compared their residents in the sample with those residing in other cities in the Haifa Bay area (*n*=`r SVIVA2_01_haifa %>% filter(close.industrial.area==0) %>% nrow()`). The data confirms that residents of these closer cities tended to perceive the air-pollution policy as more personally relevant (Cohen's D = `r relevance.areas.close.d[["estimate"]] %>% round(2) %>% abs()` [`r (relevance.areas.close.d[["conf.int"]][["upper"]]*-1) %>% round(2)`,`r (relevance.areas.close.d[["conf.int"]][["lower"]]*-1) %>% round(2)`], *t* = `r relevance.areas.close.t[["statistic"]] %>% max() %>% round(2) %>% abs()`, *p* = `r relevance.areas.close.t[["p.value"]] %>% max() %>% round(3) %>% abs()`). Second, since young children are relatively vulnerable to air-pollution effects, and since the consequences of air-pollution on infants and young children has been highlighted in media, I expect parents for young children (under 12 years) to be more engaged with this issue, although that assumption is not supported by differences in self-reported perceived personal relevance.[^footnote_post_treatment_control]

`r table.x=table.x+1`**Table `r table.x`** presents the interactions between these two variables and information manipulations. Model `r table.x`.1 presents the effect of these variables with the manipulations, and thereafter Model `r table.x`.2 adds their interactions, and finally Model `r table.x`.3 restricts the analysis to the first policy presented. All three models include control variables. The models tend to suggest that residing in the closer cities were not more likely to differentiate between the strong and weak plans, and were, to some extent, even more receptive to the emotional effect of the symbols (the interaction is significant in Model `r table.x`.2, but not in Model `r table.x`.3). As for parents for young children, the data does not show that they respond differently to the appearance of symbols or to variations in the content.[^footnote_proxies_models]

[^footnote_post_treatment_control]: In other models, not shown here, I also directly tested the interaction between the manipulations and the personal relevance subjective measures. In these models, the interactions with the symbols is insignificant and the interaction with weak information is negative and significant. However, these analysis can produce biased estimates since the perceived relevance measure is a post-manipulation variable [@gelman_2006, pp. 188-9].

[^footnote_proxies_models]: The interaction between proximity and the real symbols is positive and significant in Model `r table.x`.2, in Model `r table.x`.3  it is positive but insignificant; The interaction between parents for young children and the fake symbols is positive and significant in Model `r table.x`.2.

### 2019-07-23 paper3_revision_text.Rmd
Add files via upload

---
title: "Does personal relevance attenuate citizens' persuasion by government branding? Evidence from a survey experiment on air-pollution policy"
author: '**Saar Alon-Barkat**'
date: " "
always_allow_html: yes
output:
  word_document: default
  pdf_document:
    toc: yes
  html_document:
    theme: flatly
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
      toc_depth: 3
link-citations: yes
bibliography: phd_paper_2.bib
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa.csl
urlcolor: blue
---

@@

EPM is a national-level bureau responsible for the formulation, coordination, and execution of environmental policy in Israel. Among others, EPM is responsible for monitoring and regulating the air quality across the country. The Haifa Bay area, located in the north of Israel, has been characterized by relatively high levels of air pollution, mainly due to the Haifa Bay industrial zone, which is located in proximity of a number of urban residential areas. It is considered the largest industrial zone in the country, with concentration of petrochemical industries, including oil refineries, power plants, and chemical factories. Beginning in Mid 2000s, EPM has been developing and implementing a series of policy programs to monitor and reduce the high levels of pollution specifically in that area [@theministryofenvironmentprotection_2015; @theministryofenvironmentprotection_2015a; @theministryofenvironmentprotection_2019]. While EPM has claimed that its programs have been effective and led to a significant reduction in air-pollution levels, local residents, professionals and other public agencies have remained highly concerned about the issue, and  accordingly it remained high on the public agenda [@statecomptrollerofisrael_2019]. Specifically, the issue received much media and public attention in 2015-2016, following a series of reportings and media publications about the health consequences of the air-pollution to Haifa Bay residents. In 2015, the Ministry of Health published statistical data indicating that Haifa Bay residents experience significantly higher rates of cancer and asthma, compared to other areas of the country. The subsequent media reports specifically highlighted the conclusion that half of the cases of cancer among children in the area can be associated to air-pollution [@elroi_website_2015; @raavad_website_2015]. Thereafter, in 2016, media reporters revealed the interim results of a large epidemiological study, establishing a link between air-pollution in the Haifa Bay and certain types of cancer, as well as abnormal measurements of newborn babies and growth curves of infants.[^footnote_epidemiological_study] The media reports of that study repeatedly focused on the finding that babies in the Haifa Bay are born with "smaller heads" [@koriel_website_2016; @rinat_website_2016; @rivlin_website_2016]. Some reports also highlighted that the areas that are geographically closer to the industrial areas are found to be at higher risk for cancer [@bendavid_website_2016]. These publications, and others, caused much panic among the local population, and especially among young parents.

@@

I further examined the interactions between the manipulations and the two areas via a series of multiple linear regression models, presented in `r table.x=table.x+1`**Table `r table.x`**. I regress trust in air-pollution policy on the manipulations (Model `r table.x`.1), then I add the areas (Model `r table.x`.2) and the interaction between them (Model `r table.x`.3). The final two models are designed to test the robustness of the interaction model by controlling for individual-level covariates (Model `r table.x`.4),[^footnote_control_variables] and restricting the sample to those who saw that policy first before the recycling policy (Model `r table.x`.5). The latter analysis rules out the possibility that subjects' responses are affected by the manipulations in the first policy presented to them (a carryover effect), or by their evaluation of the it (assimilation or contrast effects). In all regression models, I use robust standard errors to adjust for heteroskedacity.[^footnote_heteroskedacity] The results of these models are largely consistent with the raw findings presented above. In Models `r table.x`.3, `r table.x`.4, `r table.x`.5 the interactions between living in the center and the real symbols are negative but not sufficiently significant, and the interactions with the weak information are insignificant. A comparison between models `r table.x`.2 and `r table.x`.3 via anova reveals that adding these interactions do not significantly improve the model fit (*F(`r anova(mod_air_1.2,mod_air_1.3) %>% tidy() %>% select(df) %>% .[2,] %>% max()`*) = `r anova(mod_air_1.2,mod_air_1.3) %>% tidy() %>% select(statistic) %>% .[2,] %>% max() %>% round(2)`, *p*= `r anova(mod_air_1.2,mod_air_1.3) %>% tidy() %>% select(p.value) %>% .[2,] %>% max() %>% round(2)`). Notice that in `r table.x`.5, the effect of the real symbols is greater, which may indicate that the impact of symbols diminishes between the first and the second time participants see them.

@@

[^footnote_recycling_model]: The main effect of the real symbols (as depicted from Model 2.1) is significant compared with the no symbols group (*Diff*=`r diff1` [`r (diff1-1.96*se1) %>% round(3)`-`r (diff1+1.96*se1) %>% round(3)`], *t*=`r (diff1/se1) %>% round(3)`). That effect amounts to `r (diff1/sd.waste) %>% round(3)` SDs [`r ((diff1-1.96*se1)/sd.waste) %>% round(3)`-`r ((diff1+1.96*se1)/sd.waste) %>% round(3)`]. Yet, the increase is not significant not with the fake symbols group (*Diff*=`r diff2` [`r (diff2-1.96*se2) %>% round(3)`-`r (diff2+1.96*se2) %>% round(3)`], *t*=`r (diff2/se2) %>% round(3)`). As shown in Model 2.3, the real symbols significantly increased Haifa Bay subjects' trust in that policy compared with the no symbols group (*Diff*=`r diff3` [`r (diff3-1.96*se3) %>% round(3)`-`r (diff3+1.96*se3) %>% round(3)`], *t*=`r (diff3/se3) %>% round(3)`), whereas the increase compared with the fake symbols group was not sufficiently significant (*Diff*=`r diff4` [`r (diff4-1.96*se4) %>% round(3)`-`r (diff4+1.96*se4) %>% round(3)`], *t*=`r (diff4/se4) %>% round(3)`). The differences between the symbols groups among the center subjects are small and insignificant. The smaller effect of the real symbols can be partly explained by the fact that I used a different celebrity campaign image for this policy, which was less familiar compared with the image presented for the air-pollution policy. The main effect of weak information (as depicted from Model 2.1) ammounts to `r (diff5/sd.waste) %>% round(3)` SDs [`r ((diff5-1.96*se5)/sd.waste) %>% round(3)`-`r ((diff5+1.96*se5)/sd.waste) %>% round(3)`]. In model 2.5, the interaction between weak information and the center area is positive, but far from significance.

@@

Next, `r table.x=table.x+1`**Table `r table.x`** presents the interaction between the relevance manipulation and the symbols and information manipulations. In Model `r table.x`.1, I regress trust in air-pollution policy on the three manipulations, then in Model `r table.x`.2 I add the interaction between the relevance manipulation and the symbols and information manipulations. Similarly to the regression tables presented above, the final two models are robust analyses where I add controls (Model `r table.x`.3) and restrict the sample to the policy presented first in order (Model `r table.x`.4). In all these models, the coefficient of the relevance treatment is negative but far from significance. The interactions between the real symbols and the relevance manipulation are negative, which indicates that among those who were primed to think about the personal relevance of the air-pollution problem, the incerase in trust due to the appearance of the symbols and those who did not were slightly smaller. However, these interaction coefficients are relatively weak (ranging between `r ((mod_relevance_1.2 %>% tidy() %>% filter(term=="SYMBOL_t1:RELEVANCE_exp") %>% select(estimate) %>% max(.) %>% abs())/sd.air.haifa) %>% round(2)`--`r ((mod_relevance_1.4 %>% tidy() %>% filter(term=="SYMBOL_t1:RELEVANCE_exp") %>% select(estimate) %>% max(.) %>% abs())/sd.air.haifa) %>% round(2)` SDs, and attenuating approximately `r ((mod_relevance_1.2 %>% tidy() %>% filter(term=="SYMBOL_t1:RELEVANCE_exp") %>% select(estimate) %>% max(.) %>% abs())/(mod_relevance_1.2 %>% tidy() %>% filter(term=="SYMBOL_t1") %>% select(estimate) %>% max(.) %>% abs())*100) %>% round(0)`--`r ((mod_relevance_1.4 %>% tidy() %>% filter(term=="SYMBOL_t1:RELEVANCE_exp") %>% select(estimate) %>% max(.) %>% abs())/(mod_relevance_1.4 %>% tidy() %>% filter(term=="SYMBOL_t1") %>% select(estimate) %>% max(.) %>% abs())*100) %>% round(0)`% of the effect in the control group) and far from significance. The interactions between the relevance treatment and weak information manipulations are also in the expected direction, but again insignificant (ranging between `r ((mod_relevance_1.4 %>% tidy() %>% filter(term=="INFORMATION_weak:RELEVANCE_exp") %>% select(estimate) %>% max(.) %>% abs())/sd.air.haifa) %>% round(2)`--`r ((mod_relevance_1.2 %>% tidy() %>% filter(term=="INFORMATION_weak:RELEVANCE_exp") %>% select(estimate) %>% max(.) %>% abs())/sd.air.haifa) %>% round(2)` SDs, and attenuating approximately `r ((mod_relevance_1.4 %>% tidy() %>% filter(term=="INFORMATION_weak:RELEVANCE_exp") %>% select(estimate) %>% max(.) %>% abs())/(mod_relevance_1.4 %>% tidy() %>% filter(term=="INFORMATION_weak") %>% select(estimate) %>% max(.) %>% abs())*100) %>% round(0)`--`r ((mod_relevance_1.2 %>% tidy() %>% filter(term=="INFORMATION_weak:RELEVANCE_exp") %>% select(estimate) %>% max(.) %>% abs())/(mod_relevance_1.2 %>% tidy() %>% filter(term=="INFORMATION_weak") %>% select(estimate) %>% max(.) %>% abs())*100) %>% round(0)`% of the effect in the control group).[^footnote_relevance_manipulation_1]

@@

Finally, as an additional test for the robustness of my findings, I further compare between Haifa Bay subjects who are more/less likely to perceive the air-pollution as personally relevant, based on two proxies. First, I rely on subjects city of residence, and its geographical proximity to the polluting industrial area. Congruently, I selected the three cities/suburbs that are closest to the industrial area - *Nesher*, *Qiryat Haiim* and *Qiryat Bialik* (*n*=`r SVIVA2_01_haifa %>% filter(close.industrial.area==1) %>% nrow()`) -- and compared their residents in the sample with those residing in other cities in the Haifa Bay area (*n*=`r SVIVA2_01_haifa %>% filter(close.industrial.area==0) %>% nrow()`). The data confirms that residents of these closer cities tended to perceive the air-pollution policy as more personally relevant (Cohen's D = `r relevance.areas.close.d[["estimate"]] %>% round(2) %>% abs()` [`r (relevance.areas.close.d[["conf.int"]][["upper"]]*-1) %>% round(2)`,`r (relevance.areas.close.d[["conf.int"]][["lower"]]*-1) %>% round(2)`], *t* = `r relevance.areas.close.t[["statistic"]] %>% max() %>% round(2) %>% abs()`, *p* = `r relevance.areas.close.t[["p.value"]] %>% max() %>% round(3) %>% abs()`). Second, since young children are relatively vulnerable to air-pollution effects, and since the consequences of air-pollution on infants and young children has been highlighted in media, I expect parents for young children (under 12 years) to be more engaged with this issue, although that assumption is not supported by differences in self-reported perceived personal relevance.[^footnote_post_treatment_control]

`r table.x=table.x+1`**Table `r table.x`** presents the interactions between these two variables and information manipulations. Model `r table.x`.1 presents the effect of these variables with the manipulations, and thereafter Model `r table.x`.2 adds their interactions, and finally Model `r table.x`.3 restricts the analysis to the first policy presented. All three models include control variables. The models tend to suggest that residing in the closer cities were not more likely to differentiate between the strong and weak plans, and were, to some extent, even more receptive to the emotional effect of the symbols (the interaction is significant in Model `r table.x`.2, but not in Model `r table.x`.3). As for parents for young children, the data does not show that they respond differently to the appearance of symbols or to variations in the content.[^footnote_proxies_models]

[^footnote_post_treatment_control]: In other models, not shown here, I also directly tested the interaction between the manipulations and the personal relevance subjective measures. In these models, the interactions with the symbols is insignificant and the interaction with weak information is negative and significant. However, these analysis can produce biased estimates since the perceived relevance measure is a post-manipulation variable [@gelman_2006, pp. 188-9].

[^footnote_proxies_models]: The interaction between proximity and the real symbols is positive and significant in Model `r table.x`.2, in Model `r table.x`.3  it is positive but insignificant; The interaction between parents for young children and the fake symbols is positive and significant in Model `r table.x`.2.

@@

This study contributes to the growing public-administration research on government branding and its potency to influence and possibly manipulate citizens' perceptions of government organizations and their policies and services. I explored whether persuasion through branding and symbolic communication can also ocure in policy matters that have significant, clear implications for citizens' well-being. I examined the hypothesis, derived from social psychology theory, that persuasion via symbols is attenuated the more citizens perceive the communication as personally relevant. I put this hypothesis to rigorous empirical testing using a randomized survey experiment focused on air-pollution policy, utilizing a natural variation in the perceived personal relevance between citizens residing in a polluted area and others, residing elsewhere. To the best of my knowledge, this is the first study to explore, theoretically and empirically, the moderating role of perceived personal relevance on citizens' responses to government branding, and more generally to government public communications.

@@

How can we reconcile the findings of this study with the proposed theory? One plausible explanation is that despite the relatively high involvement of Haifa Bay residents, they were still not motivated enough to rely only on the central route and disregard the peripheral cues. If this explanation holds, it entails that citizens may resist the emotional effect of symbols only in rare circumstances, which are characterized by extremely high levels of personal relevance, such as a policy that has immediate and tangible implications on citizens well-being. Indeed, social psychology ELM scholars Petty and Cacioppo have acknowledged that the high and low personal relevance conditions generated in their experimental settings were deliberately created for theory testing purposes, and can be rarely found in the world outside the laboratory [@petty_1986, p. 206].

How can we explain the finding that the positive effect of the real symbols was stronger to some extent among Haifa Bay residents? Given that this pattern is also evident for the recycling policy, it is unlikely that it is linked to differences in perceived personal relevance. Rather, I tentatively suggest two other possible explanations for that. First, these patterns are resulted by the lower baseline of Haifa Bay residents' trust in EPM policies. Accordingly, it could be the case that symbolic elements are more effective in shifting people from low to medium levels of trust (i.e. as in the case of Haifa Bay subjects), than shifting them from medium to high levels (i.e. as in the case of center subjects), indicating a ceiling effect. In other words, symbols may be more effective in reducing citizens' distrust, then increasing their trust.[^footnote_ceiling_effect] A second explanation could be that EPM symbols attenuated the effect of Haifa Bay residents' prior beliefs on their judgements of the policy plans, an effect that can be related to motivated reasoning [@taber_2006; @christensen_2018; @baekgaard_2016]. It can be theorized that symbols attenuate motivated reasoning, when symbolic associations are not strongly related to identities of groups having strong prior positions on the policy or the organization (as in the case of the EPM symbols and the Haifa Bay local residents). On the other hand, when symbolic associations do activate such group identities, they are likely to exacerbate partisan motivated reasoning in that regard, and polarization [@teodoro_2018]. Certainly, we must be cautious about these tentative explanations, which may be tested by future studies.

@@

The results of this study, if generalizable, have significant implications. They suggest that citizens are susceptible to persuasion by government branding and symbolic communication, even with regard to salient, consested policy issues that have significant consequences for their own lives. Accordingly, these findings suggest that the boundaries of persuasion via branding are much wider than expected. This entails that public organizations can affect citizens in a broad array of policy domains and contexts by investing in symbolic communications. Alongside the benefits of this instrument in mitigating citizens' distrust, it also has the potency to yield a distorted, emotive, overoptimistic view of government organizations and their policies, which may undermine basic democratic accountability mechanisms. For good or for ill, citizens are affected by the emotional connections and associations of symbols.

>>>>

---
title: "Does personal relevance attenuate citizens' persuasion by government branding? Evidence from a survey experiment on air-pollution policy"
author: '**Saar Alon-Barkat**'
date: " "
always_allow_html: yes
output:
  html_document:
    theme: flatly
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
      toc_depth: 3
  pdf_document:
    toc: yes
  word_document: default
link-citations: yes
bibliography: phd_paper_2.bib
csl: https://raw.githubusercontent.com/citation-style-language/styles/master/apa.csl
urlcolor: blue
---

@@

EPM is a national-level bureau responsible for the formulation, coordination, and execution of environmental policy in Israel. Among others, EPM is responsible for monitoring and regulating the air quality across the country. The Haifa Bay area, located in the north of Israel, has been characterized by relatively high levels of air pollution, mainly due to the Haifa Bay industrial zone, which is located in proximity of a number of urban residential areas. It is considered the largest industrial zone in the country, with concentration of petrochemical industries, including oil refineries, power plants, and chemical factories. Beginning in Mid 2000s, EPM has been developing and implementing a series of policy programs to monitor and reduce the high levels of pollution specifically in that area [@theministryofenvironmentprotection_2015; @theministryofenvironmentprotection_2015a; @theministryofenvironmentprotection_2019]. While EPM has claimed that its programs have been effective and led to a significant reduction in air-pollution levels, local residents, professionals and other public agencies have remained highly concerned about the issue, and  accordingly it remained high on the public agenda [@statecomptrollerofisrael_2019]. Specifically, the issue received much media and public attention in 2015-2016, following a series of reportings and media publications about the health consequences of the air-pollution to Haifa Bay residents. In 2015, the Ministry of Health published statistical data indicating that Haifa Bay residents experience significantly higher rates of cancer and asthma, compared to other areas of the country. Thereafter, in 2016, media reporters revealed the interim results of a large epidemiological study, establishing a link between air-pollution in the Haifa Bay and certain types of cancer, as well as abnormal measurements of newborn babies and growth curves of infants.[^footnote_epidemiological_study] The media reports of that study repeatedly highlighted a particular finding that babies in the Haifa Bay are born with "smaller heads" [@koriel_website_2016; @rinat_website_2016; @rivlin_website_2016]. These publications, and others, caused much panic among the local population, and especially among young parents.

@@

I further examined the interactions between the manipulations and the two areas via a series of multiple linear regression models, presented in `r table.x=table.x+1`**Table `r table.x`**. I regress trust in air-pollution policy on the manipulations (Model `r table.x`.1), then I add the areas (Model `r table.x`.2) and the interaction between them (Model `r table.x`.3). In Model `r table.x`.4, I test the robustness of the interaction model by controlling for individual-level covariates.[^footnote_control_variables] In all regression models, I use robust standard errors to adjust for heteroskedacity.[^footnote_heteroskedacity] The results of these models are largely consistent with the raw findings presented above. In Models `r table.x`.3, `r table.x`.4 the interactions between living in the center and the real symbols are negative but not sufficiently significant, and the interactions with the weak information are insignificant. A comparison between models `r table.x`.2 and `r table.x`.3 via anova reveals that adding these interactions do not significantly improve the model fit (*F(`r anova(mod_air_1.2,mod_air_1.3) %>% tidy() %>% select(df) %>% .[2,] %>% max()`*) = `r anova(mod_air_1.2,mod_air_1.3) %>% tidy() %>% select(statistic) %>% .[2,] %>% max() %>% round(2)`, *p*= `r anova(mod_air_1.2,mod_air_1.3) %>% tidy() %>% select(p.value) %>% .[2,] %>% max() %>% round(2)`).

@@

[^footnote_recycling_model]: The main effect of the real symbols (as depicted from Model 2.1) is significant compared with the no symbols group (*Diff*=`r diff1` [`r (diff1-1.96*se1) %>% round(3)`-`r (diff1+1.96*se1) %>% round(3)`], *t*=`r (diff1/se1) %>% round(3)`). Yet, the increase is not significant not with the fake symbols group (*Diff*=`r diff2` [`r (diff2-1.96*se2) %>% round(3)`-`r (diff2+1.96*se2) %>% round(3)`], *t*=`r (diff2/se2) %>% round(3)`). As shown in Model 2.3, the real symbols significantly increased Haifa Bay subjects' trust in that policy compared with the no symbols group (*Diff*=`r diff3` [`r (diff3-1.96*se3) %>% round(3)`-`r (diff3+1.96*se3) %>% round(3)`], *t*=`r (diff3/se3) %>% round(3)`), whereas the increase compared with the fake symbols group was not sufficiently significant (*Diff*=`r diff4` [`r (diff4-1.96*se4) %>% round(3)`-`r (diff4+1.96*se4) %>% round(3)`], *t*=`r (diff4/se4) %>% round(3)`). The differences between the symbols groups among the center subjects are small and insignificant. The weaker effect of the real symbols can be partly explained by the fact that I used a different celebrity campaign image for this policy, which was less familiar compared with the image presented for the air-pollution policy. The main effect of weak information (as depicted from Model 2.1) ammounts to `r (diff5/sd.waste) %>% round(3)` SDs [`r ((diff5-1.96*se5)/sd.waste) %>% round(3)`-`r ((diff5+1.96*se5)/sd.waste) %>% round(3)`]. In model 2.5, the interaction between weak information and the center area is positive, but far from significance.

@@

Next, `r table.x=table.x+1`**Table `r table.x`** presents the interaction between the relevance manipulation and the symbols and information manipulations. In Model `r table.x`.1, I regress trust in air-pollution policy on the three manipulations, then in Model `r table.x`.2 I add the interaction between the relevance manipulation and the symbols and information manipulations, and in Model `r table.x`.3 I add controls. In all these models, the coefficient of the relevance treatment is negative but far from significance. The interactions between the real symbols and the relevance manipulation are negative, which indicates that among those who were primed to think about the personal relevance of the air-pollution problem, the incerase in trust due to the appearance of the symbols and those who did not were slightly smaller. However, these interaction coefficients are relatively weak and far from significance (ranging between `r ((mod_relevance_1.2 %>% tidy() %>% filter(term=="SYMBOL_t1:RELEVANCE_exp") %>% select(estimate) %>% max(.) %>% abs())/sd.air.haifa) %>% round(2)`--`r ((mod_relevance_1.3 %>% tidy() %>% filter(term=="SYMBOL_t1:RELEVANCE_exp") %>% select(estimate) %>% max(.) %>% abs())/sd.air.haifa) %>% round(2)` SDs, and attenuating approximately `r ((mod_relevance_1.2 %>% tidy() %>% filter(term=="SYMBOL_t1:RELEVANCE_exp") %>% select(estimate) %>% max(.) %>% abs())/(mod_relevance_1.2 %>% tidy() %>% filter(term=="SYMBOL_t1") %>% select(estimate) %>% max(.) %>% abs())*100) %>% round(0)`--`r ((mod_relevance_1.3 %>% tidy() %>% filter(term=="SYMBOL_t1:RELEVANCE_exp") %>% select(estimate) %>% max(.) %>% abs())/(mod_relevance_1.3 %>% tidy() %>% filter(term=="SYMBOL_t1") %>% select(estimate) %>% max(.) %>% abs())*100) %>% round(0)`% of the effect in the control group). The interactions between the relevance treatment and weak information manipulations are also in the expected direction, but again insignificant (ranging between `r ((mod_relevance_1.3 %>% tidy() %>% filter(term=="INFORMATION_weak:RELEVANCE_exp") %>% select(estimate) %>% max(.) %>% abs())/sd.air.haifa) %>% round(2)`--`r ((mod_relevance_1.2 %>% tidy() %>% filter(term=="INFORMATION_weak:RELEVANCE_exp") %>% select(estimate) %>% max(.) %>% abs())/sd.air.haifa) %>% round(2)` SDs, and attenuating approximately `r ((mod_relevance_1.3 %>% tidy() %>% filter(term=="INFORMATION_weak:RELEVANCE_exp") %>% select(estimate) %>% max(.) %>% abs())/(mod_relevance_1.3 %>% tidy() %>% filter(term=="INFORMATION_weak") %>% select(estimate) %>% max(.) %>% abs())*100) %>% round(0)`--`r ((mod_relevance_1.2 %>% tidy() %>% filter(term=="INFORMATION_weak:RELEVANCE_exp") %>% select(estimate) %>% max(.) %>% abs())/(mod_relevance_1.2 %>% tidy() %>% filter(term=="INFORMATION_weak") %>% select(estimate) %>% max(.) %>% abs())*100) %>% round(0)`% of the effect in the control group).[^footnote_relevance_manipulation_1]

@@

I further explored the robustness of these findings through additional analyses, reported in the supplementary appendix. *First*, I replicate the main models while restricting the analyses to the first policy seen by the participants. This enables to rule out the possibility that subjects' responses are affected by the manipulations in the first policy presented to them (a carryover effect), or by their evaluation of the it (assimilation or contrast effects). *Second*, I restrict the analyses to Haifa Bay residents, and compare between those who are more/less likely to perceive the air-pollution as personally relevant, based on two proxies: (a) Subjects city of residence, and its geographical proximity to the polluting industrial area; (b) parents for young children. *Third*, I replicated the analyses on the unfiltered sample. These additional analyses are consistent with the analyses presented in the paper, and do not provide evidence that greater levels of perceived personal relevance attenuates the effect of symbols.

@@

This study contributes to the growing public-administration research on government branding and its potency to influence and possibly manipulate citizens' perceptions of government organizations and their policies and services. I explored whether persuasion through branding and symbolic communication can also occur in policy matters that have significant, clear implications for citizens' well-being. I examined the hypothesis, derived from social psychology theory, that persuasion via symbols is attenuated the more citizens perceive the communication as personally relevant. I put this hypothesis to rigorous empirical testing using a randomized survey experiment focused on air-pollution policy, utilizing a natural variation in the perceived personal relevance between citizens residing in a polluted area and others, residing elsewhere. To the best of my knowledge, this is the first study to explore, theoretically and empirically, the moderating role of perceived personal relevance on citizens' responses to government branding, and to government public communications at large.

@@

How can we reconcile the findings of this study with the proposed theory? One plausible explanation is that despite the relatively high involvement of Haifa Bay residents, they were still not motivated enough to rely only on the central route and disregard the peripheral cues. If this explanation holds, it entails that citizens may resist the emotional effect of symbols only in rare circumstances, which are characterized by extremely high levels of personal relevance -- policies that have immediate, tangible and unambiguous consequences for citizens well-being. Indeed, social psychology ELM scholars Petty and Cacioppo have acknowledged that the high and low personal relevance conditions generated in their experimental settings were deliberately created for theory testing purposes, and can be rarely found in the world outside the laboratory [@petty_1986, p. 206].

How can we explain the finding that the positive effect of the real symbols was stronger to some extent among Haifa Bay residents? Given that this pattern is also evident for the recycling policy, it is unlikely that it is linked to differences in perceived personal relevance. Rather, I tentatively suggest two other possible explanations for that. First, these patterns could be the result of the higher intercept of center residents' trust in EPM policies, or a ceiling effect. More generally, it could be the case that symbolic elements are more effective in shifting people from low to medium levels of trust (i.e. as in the case of Haifa Bay subjects), than shifting them from medium to high levels (i.e. as in the case of center subjects). In other words, symbols may be more effective in reducing citizens' distrust, then increasing their trust.[^footnote_ceiling_effect] A second explanation could be that EPM symbols attenuated the effect of Haifa Bay residents' prior beliefs on their judgments of the policy plans, an effect that can be related to motivated reasoning [@taber_2006; @christensen_2018; @baekgaard_2016]. It can be theorized that symbols attenuate motivated reasoning, when symbolic associations are not strongly related to identities of specific groups having strong prior positions on the policy or the organization (as in the case of the EPM symbols and the Haifa Bay locals). On the other hand, when symbolic associations do activate such group identities, they are likely to exacerbate partisan motivated reasoning in that regard, and polarization [@teodoro_2018]. For instance, it is possible that EPM "green" symbols enhance partisan motivated reasoning among groups divided on their support for environmental policy. Certainly, we must be cautious about these tentative explanations, which may be tested by future studies.

@@

The results of this study, if generalizable, have significant implications. They suggest that citizens are susceptible to persuasion by government branding and symbolic communication, even with regard to salient, contested policy issues that have significant consequences for their own lives. Accordingly, these findings suggest that the boundaries of persuasion via branding are much wider than expected. This entails that public organizations can affect citizens in a broad array of policy domains and contexts by investing in symbolic communications. Alongside the benefits of this instrument in mitigating citizens' distrust, it also has the potency to yield a distorted, emotive, overoptimistic view of government organizations and their policies, which may undermine basic democratic accountability mechanisms. For good or for ill, citizens are affected by the emotional connections and associations of symbols.
